Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-09 paper/geometry-twocolumn.tex
Tidy up the two outstanding FIXMEs

Invariance under change of origin is the key feature of affine
geometry, whose central operation is the affine combination of points:
$\lambda_1\vec{v_1} + \lambda_2\vec{v_2}$, where $\lambda_1 +
\lambda_2 = 1$.  Geometrically, this can be interpreted as
interpolation between the points represented by $\vec{v_1}$ and
$\vec{v_2}$.  We add affine combination of points to our calculus as
follows: % typing and intended denotation:for
\begin{displaymath}
  \begin{array}{l}
    \mathrm{affComb} :\forall t \mathord:
    \SynTransl{2}.\ \tyPrim{vec}{t} \to \tyPrimNm{real} \to
    \tyPrim{vec}{t} \to \tyPrim{vec}{t} \\ 
    \sem{\mathrm{affComb}}\ \vec{v_1}\ r\ \vec{v_2} = (1-r)\vec{v_1} +
    r\vec{v_2} 
\end{array}
\end{displaymath}
It can be verified by hand that %the intended denotation
$\sem{\mathrm{affComb}}$ has the invariance property specified by its
type. {\bf Vague. Say which invariance property.}
% By defining the primitive function $\mathrm{affComb}$ to take a
% single $\tyPrimNm{real}$ parameter $t$, we can easily ensure that we
% are taking the affine combination of two representatives of points,
% and not just an arbitrary linear combination.

@@

Indexing types by scaling factors brings us to the original
inspiration for the current work: Kennedy's interpretation of his
units of measure type system via scaling invariance
\cite{kennedy97relational}. Kennedy shows how interpreting types in
terms of scaling invariance brings the techniques of dimensional
analysis \fixme{ref} to bear on programming. The types of the real
number arithmetic operations in \autoref{fig:real-ops} are exactly the
types Kennedy assigns in his units of measure system, except for that
of the absolute value operation. Semantically, our type indexes by
%our assigned type is that, semantically, we are indexing by
non-zero scaling factors, whereas Kennedy's indexes by strictly
positive ones. We examine the typing of the absolute value operation
further in \autoref{sec:types-indexed-abelian-groups}.

>>>>

Invariance under change of origin is the key feature of affine
geometry, whose central operation is the affine combination of points:
$\lambda_1\vec{v_1} + \lambda_2\vec{v_2}$, where $\lambda_1 +
\lambda_2 = 1$.  Geometrically, this can be interpreted as
interpolation between the points represented by $\vec{v_1}$ and
$\vec{v_2}$.  We add affine combination of points to our calculus as
follows: % typing and intended denotation:for
\begin{displaymath}
  \begin{array}{l}
    \mathrm{affComb} :\forall t \mathord:
    \SynTransl{2}.\ \tyPrim{vec}{t} \to \tyPrimNm{real} \to
    \tyPrim{vec}{t} \to \tyPrim{vec}{t} \\ 
    \sem{\mathrm{affComb}}\ \vec{v_1}\ r\ \vec{v_2} = (1-r)\vec{v_1} +
    r\vec{v_2} 
\end{array}
\end{displaymath}
It can be verified by hand that %the intended denotation
$\sem{\mathrm{affComb}}$ is invariant under all changes of origin, as
dictated by its type.
% By defining the primitive function $\mathrm{affComb}$ to take a
% single $\tyPrimNm{real}$ parameter $t$, we can easily ensure that we
% are taking the affine combination of two representatives of points,
% and not just an arbitrary linear combination.

@@

Indexing types by scaling factors brings us to the original
inspiration for the current work: Kennedy's interpretation of his
units of measure type system via scaling invariance
\cite{kennedy97relational}. Kennedy shows how interpreting types in
terms of scaling invariance brings the techniques of dimensional
analysis to bear on programming. The types of the real number
arithmetic operations in \autoref{fig:real-ops} are exactly the types
Kennedy assigns in his units of measure system, except for that of the
absolute value operation. Semantically, our type indexes by
%our assigned type is that, semantically, we are indexing by
non-zero scaling factors, whereas Kennedy's indexes by strictly
positive ones. We examine the typing of the absolute value operation
further in \autoref{sec:types-indexed-abelian-groups}.
